Introduction to Lattices Attacks\\``Cryptography''

### 2021-06-04 paper.tex
ROCA explanation

\section{Problems}

@@

We start introducing the main ideas of Coppersmith's attack \cite{coppersmith96} and the math behind it.
The attack can be used against \textbf{RSA} when a small $e$ is used. In 2017 the attack was used to exploit
\href{https://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2017-15361}{CVE-2017-15361} also named \textbf{ROCA}
(Return Of Coppersmith's Attack) \cite{roca}.

@@

\[
B = 
\begin{pmatrix}
    X^3 & f_2X^2 & f_1X & f_0 \\
    0 & NX^2 & 0 & 0 \\
    0 & 0 & NX & 0 \\
    0 & 0 & 0 & N
\end{pmatrix}
\]

The rows of the matrix correspond to the coefficient vectors of the polynomials $f(x), Nx^2, Nx$ and $N$, furthermore we know
that each polynomials will be 0 modulo $N$ if evaluated at $x = m$.
We applied Howgrave-Graham with $m=1$ (the $N^m$ parameter not the message).\\
With this lattice construction every vector is of the form $v = (v_1X^3, v_2X^2, v_1X, v_0)$, because any
integer linear combination of the vector of the lattice will keep the bound $X^i$ for $i=0,..,\dim(B)-1$.

@@

Knowing the most significat bits of $p$ can be used to factor the modulus $N$ ref TODO coppersmith.

Let $a = 2^lb$ where $b$ are the bits known of $p$, then $p = a + r$ for some small value of $r$.

Using as lattice basis

@@

And applying LLL we can find $r$.

>>>>

\section{Lattice Problems}

@@

We start introducing the main ideas of Coppersmith's attack \cite{coppersmith96} and the math behind it.
The attack can be used against \textbf{RSA} when a small $e$ is used, but also
to factorize the modulus. In 2017 the attack was used to exploit
\href{https://cve.mitre.org/cgi-bin/cvename.cgi?name=CVE-2017-15361}{CVE-2017-15361} also named \textbf{ROCA}
(Return Of Coppersmith's Attack) \cite{roca}.

@@

\[
B = 
\label{mat:lattice_rsa}
\begin{pmatrix}
    X^3 & f_2X^2 & f_1X & f_0 \\
    0 & NX^2 & 0 & 0 \\
    0 & 0 & NX & 0 \\
    0 & 0 & 0 & N
\end{pmatrix}
\]

The rows of the matrix correspond to the coefficient vectors of the polynomials $f(x), Nx^2, Nx$ and $N$, furthermore we know
that each polynomials will be 0 modulo $N$ if evaluated at $x = m$.
We applied Howgrave-Graham with $m=1$ (the $N^m$ parameter not the message).\\
With this lattice construction every vector is of the form $v = (v_3X^3, v_2X^2, v_1X, v_0)$, because any
integer linear combination of the vector of the lattice will keep the bound $X^i$ for $i=0,..,\dim(B)-1$.

@@

Lattices can also be used to factor a modulus $N$ knowing the most significant bits of one of the factor.
Let $a = 2^lb$ where $b$ are the bits known of $p$, then $p = a + r$ for some small value of $r$.
We can rewrite the problem to find the root $r$ of $f(x) = a + x \mod p$ such that $f(r) = p \equiv 0 \mod p$.
Suppose that we know the bound of $r$ such that $|r| < X$ for some bound $X$.

We can construct the following lattices

@@

Notice that the rows correspond to the polynomials $x(x + a)$, $(x + a)$, $N$ and that each of these polynomials will
be $0 \mod p$ when $x = r$. As before \ref{mat:lattice_rsa}, every polynomial is scaled by $X$ which is the upperbound.

We need to verify that the shortest vector that we can with \textbf{LLL} will be less than $p$ because if $g(x) < p$,
then by construction we can compute $g(r) = 0$ over the integers.
In this case we have

\[
    (detB)^{\frac{1}{\dim L}} = (X^3N)^{\frac{1}{3}} < p
\]

Solving for $X$ we obtain $X < p^{\frac{1}{3}}$, so this method works for every $|r| < p^{\frac{1}{3}}$.

To find $r$ we just need to apply \textbf{LLL} and take the first vector (shortest) that will be of the form $v = (v_2X^2, v_1X, v_0)$.

The last step is to construct $g(x) = v_2x^2 + v_1x + v_0$ (just divides every coefficient by $X^i$)
and computes the roots over the integers to find $r$. We implemented the attack against a modulus
of \texbtf{RSA}-$1024$ knowing the $160$ most significant bits of $p$. The attack can be expanded
until $X < p^{\frac{1}{2}}$ with a bigger lattice.

\subsection{Real World case study: ROCA}

\textbf{ROCA}\cite{roca} is a vulnerability discovered in the 2017 in a software library \textit{RSALib} in the generation
of the public modulus for \textbf{RSA}. The attack is believed to affects millions of smart cards. The vulnerbality
resides in the generation of prime numbers, since embedded devices don't have lots of computational resources a fast way
was needed. Every prime generated has the following structure:

\[
    p = kM + (g^a \mod M)
    \label{eq:roca}
\]

Where $k, a$ are random values different for every prime (unkown), $g = 65537$ and $M$ is known and it's a \textit{primorial} number.
A \textit{primorial} number $M$ is the product of the the first $n$ successive primes

\[
    M = P_n# = \prod_{i=1}^{n} P_i
\]

where $n$ is related to key size. In the case of \textit{RSALib} the primorial number used had similar dimension to $p$,
and the value $a, k$ didn't have much entropy. The modulus is

\[
    N = (kM + g^{a}\mod M)(lM + g^{b}\mod M)
\]

for $a, b, k, l \in \Z$. It's equivalent to

\[
    N \equiv g^{a + b} \equiv g^{c} \mod M
\]

for some $c \in \Z$ since all the multiples of $kM$ or $lM$ are $0 \mod M$.

It's possible to compute the discrete logarithm of $g^c \mod M$ since the size of $|M|$ is smooth with
Pohlig-Hellman algorithm \cite{pohlig}. With $c$ we can set a bound to the value of $a$ since

\[
    \frac{c}{2} \le a \le \frac{c + ord}{2}
\]

where $ord$ is the multiplicative order of the generator $g \in Z_M$.

Once we know the bound of $a$ we can try to find the value of $k$ with the Coppersmith's method \cite{coppersmith96} for every possible $a$.

The concept of the algorithm is the following:

\begin{algorithm}[H]
    \vspace*{5px}
    \textbf{Input: } $N, M, g, ord, X$ \; \\
    $c = \log_{g}N \mod M$ \;\\
    \For{$a \gets (\frac{c}{2}, \frac{c + ord}{2})$}{
        \vspace*{5px}
        $f(x) = xM + (g^a \mod M)$ \; \\
        $g(x) = \frac{1}{M}f(x)$ [make monic, the roots are the same] \; \\
        $k = $ coppersmith($g(x), N, X$) [find k] \; \\ 
        $p = kM + (g^a \mod M)$  [candidate] \; \\
        \If{$N \mod p = 0$}{
            return $p$ \;
        }
    }
\end{algorithm}

In practice however the bound of $a$ is too big, so the authors found a way the problem to a smaller $M'$ since the
bitsize of $M$ are more than sufficient for making the attack works with Coppersmith's algorithm. Their optimization
find a new $M'$ such that

\begin{itemize}
    \item $(p, q)$ are still of the form \ref{eq:roca}, so $M'$ must be a divisor of $M$.
    \item Coppersmith's algorithm will find $k'$ for the correct guess of $a'$.
    \item Find a small order of $g \in \Z_M'$ to make $a'$ smaller.
\end{itemize}

We were able to reproduce the attack in sagemath against \textbf{RSA}-512, and we used as $M'$ the value
that Bruno Produit \cite{rocaopt} made available in his thesis and David Wong's implementation of Coppersmith's attack \cite{wong2015}.
